# In-Class Programming Activity, Day 1

# Hannah Sonsalla

## Task 1

Load mosaic package into R.

```{r}
library(mosaic)
```

## Task2

I created an object named task2 that contains a character string like "Today is Tue Sep 20 07:16.11 2016" that reflects the correct date on any day in the future that the document is compiled.
```{r}
task2 <- paste("Today is", date()) 
task2
```


## Task 3


```{r}
galtonData <- mosaicData::Galton

task3a <- names(galtonData)
task3a

task3b <- nrow(galtonData)
task3b

task3c <- mean(galtonData$height)
task3c
```

The variable names of the Galton data frame are family, father, mother, sex, height, and nkids.  There are 898 cases in the Galton data frame.  The mean of height in Galton is 66.76069 inches.

## Task 4
I created a matrix using the rbind function to bind together 2 rows.
```{r}
task4 <- rbind(c(1,2,3), c(4,5,6))
task4
```

## Task 5
```{r}
task5x = c(runif(1000)) 
task5y = c(runif(1000))
task5pia = sqrt((task5x^2) + (task5y^2))
task5pi = (sum(task5pia <= 1) / 1000) * 4
task5pi
```

The variable task5pi estimates pi from a "Monte Carlo" simulation.



## Test statements

```{r}
scoreActivity::score253(1)
```