# In-Class Programming Activity, Day 2

### *Hannah Sonsalla*

## Task 1

I created an object called auto_file_name that contains the full path name to the Auto.data file on my system.
```{r}
#download.file("http://www-bcf.usc.edu/~gareth/ISL/Auto.csv", destfile="Auto.csv")

auto_file_name <- "/home/local/MAC/hsonsall/Math-253-Assignment/Auto.csv"
```

## Task 2

I created an object called Auto using the read.csv().
```{r}
Auto <- read.csv(auto_file_name)
```

## Task 3

I created an object task3 that holds a summary of the horsepower variable.
```{r}
task3 <- summary(Auto$horsepower)
```

## Task 4

I installed the ISLR package.
```{r}
library(ISLR)
```

## Task 5
I used indexing operations described to create the following objects:

task5top - the subset of Auto consisting of the first 5 rows in the first 3 columns.
task5bottom - the subset of Auto consisting of the last 5 rows in the last 3 columns.

I made an object called task5rowcolumn to hold the number of rows and columns of the Auto data.

```{r}
task5top <- Auto[1:5, 1:3]
task5rowcolumn <- dim(Auto)
task5rowcolumn
task5bottom <- Auto[393:397, 7:9]
```

## Test statements

```{r}
scoreActivity::score253(2)
```