# In-Class Programming Activity, Day 9
### *Hannah Sonsalla*

## Task 1
Create deck of cards.  The first two digits represent the rank (with 11 for J, 12 for Q, and so on). The last digit represents the suit: 1, 2, 3, 4. For instance, the card 123 is the queen of hearts, since 12 in the first two digits stands for Q and 3 in the last digit stands for hearts.
```{r}
poker_deck <- c(outer((10 * 2:14), 1:4, '+'))
poker_deck
```

Functions suits and ranks.
```{r}
suits = function(cards){
  return(cards%%10)
}

ranks = function(cards){
  return(cards%/%10)
} 

#Testing suits() and ranks() functions
#suits(c(123, 134))
#ranks(c(123, 134))
```

Test hands:
```{r}
#Royal Flush
test_hand_1 = c(102,122,132,112,142)
#Straight Flush
test_hand_2 = c(42,52,72,82,62)
#Four of a kind
test_hand_3 = c(42,43,44,41,82)
#Full house
test_hand_4 = c(42,83,44,41,82)
#Flush
test_hand_5 = c(102,82,42,72,132)
#Straight
test_hand_6 = c(102,83,71,93,62)
#Three of a kind
test_hand_7 = c(134, 131, 52, 73, 133)
#Two pair
test_hand_8 = c(102, 83, 43, 84, 101)
#Pair
test_hand_9 = c(54, 72, 141, 102, 53)

#No combo test hands
test_hand_10 = c(21, 34, 142, 113, 62)
test_hand_11 = c(91, 124, 33, 54, 102)

```

Royal flush (five cards of the same suit with ranks 10, ace, queen, king, ace):
```{r}
is_royal_flush = function(five_cards){
  card_suits = table(suits(five_cards))
  min_suit = min(card_suits)
  max_suit = max(card_suits)
  card_ranks = ranks(five_cards)
  
  ifelse((min(card_ranks) == 10) && max(diff(sort(card_ranks))) == 1 && (min_suit == max_suit), TRUE, FALSE)

}

#Tests for royal_flush
is_royal_flush(test_hand_1)

#False 
is_royal_flush(test_hand_10)
```

Straight flush (five cards of the same suit in consecutive order):
```{r}
is_straight_flush = function(five_cards){
  card_suits = table(suits(five_cards))
  min_suit = min(card_suits)
  max_suit = max(card_suits)
  card_ranks = ranks(five_cards)

  ifelse(min_suit == max_suit && max(diff(sort(card_ranks))) == 1 && is_royal_flush(five_cards) == FALSE, TRUE, FALSE)
}

# Tests for straight_flush
is_straight_flush(test_hand_2)

# Not royal flush
is_straight_flush(test_hand_1)

# False
is_straight_flush(test_hand_9)
```

Four of a kind (four cards of the same value):
```{r}
is_four_of_a_kind = function(five_cards) {
  card_ranks = table(ranks(five_cards))
  ifelse(max(card_ranks) == 4, TRUE, FALSE)
}
#Tests for four of a kind
is_four_of_a_kind(test_hand_3)

#False
is_four_of_a_kind(test_hand_11)
```

Full house(Three of a kind and the other two cards forming a pair):
```{r}
is_full_house = function(five_cards) {
  card_ranks = table(ranks(five_cards))
  ifelse(max(card_ranks) == 3 && min(card_ranks) == 2, TRUE, FALSE)
}

#Tests for full house
is_full_house(test_hand_4)

# False
is_full_house(test_hand_10)
```

Flush (five cards of the same suit):
```{r}
is_flush = function(five_cards){
  card_suits = table(suits(five_cards))
  min_suit = min(card_suits)
  max_suit = max(card_suits)
  ifelse(min_suit == max_suit && is_royal_flush(five_cards) == FALSE && is_straight_flush(five_cards) == FALSE, TRUE, FALSE)
}

#Tests for flush
is_flush(test_hand_5)

# Royal Flush
is_flush(test_hand_1)
#Straight Flush
is_flush(test_hand_2)
#False
is_flush(test_hand_11)
```

Straight (five consecutively numbered cards):
```{r}
is_straight = function(five_cards) {
  card_ranks = ranks(five_cards)
  ifelse(is_royal_flush(five_cards) == FALSE && is_straight_flush(five_cards) == FALSE && max(diff(sort(card_ranks))) == 1, TRUE, FALSE)
}

#Tests for straight
is_straight(test_hand_6)

# Royal Flush
is_straight(test_hand_1)
# Straight Flush
is_straight(test_hand_2)
# False
is_straight(test_hand_11)
```

Three of a kind (three cards same rank):
```{r}
is_three_of_a_kind = function(five_cards){
  card_ranks = table(ranks(five_cards))
  ifelse(max(card_ranks) == 3 && is_full_house(five_cards) == FALSE, TRUE, FALSE)
}

# Tests for three of a kind
is_three_of_a_kind(test_hand_7)

# Royal Flush
is_three_of_a_kind(test_hand_1)
# Straight Flush
is_three_of_a_kind(test_hand_2)
# Four of a Kind
is_three_of_a_kind(test_hand_3)
# Full house
is_three_of_a_kind(test_hand_4)
# Flush
is_three_of_a_kind(test_hand_5)
# Straight
is_three_of_a_kind(test_hand_6)
# False
is_three_of_a_kind(test_hand_10)
```

Two pair:
```{r}
is_two_pair = function(five_cards){
  card_ranks = table(ranks(five_cards))
  ifelse(min(card_ranks) == 1 && max(card_ranks) == 2 && dim(card_ranks) == 3, TRUE, FALSE)
}

# Tests for two pair
is_two_pair(test_hand_8)

#False
is_two_pair(test_hand_11)

```

Pair (two cards of same rank):
```{r}
is_pair = function(five_cards) {
  card_ranks = table(ranks(five_cards))
  ifelse(max(card_ranks) == 2 && is_two_pair(five_cards) == FALSE && is_three_of_a_kind(five_cards) == FALSE && is_full_house(five_cards) == FALSE && is_four_of_a_kind(five_cards) == FALSE, TRUE, FALSE)
}

# Tests for pair

#Pair
is_pair(test_hand_9)

#Four of a kind
is_pair(test_hand_3)
#Full house
is_pair(test_hand_4)
#Three of a kind
is_pair(test_hand_7)
# Two pair
is_pair(test_hand_8)

# False
is_pair(test_hand_10)
```

Before draw function
```{r}
before_draw = function(x) {
  five_cards <- sample(poker_deck, 5)
  if(is_royal_flush(five_cards)) {
    return ("royal flush")
  } 
  else if (is_straight_flush(five_cards)) {
    return ("straight flush")
  } 
  else if (is_four_of_a_kind(five_cards)) {
    return ("four of a kind")
  }
  else if (is_full_house(five_cards)) {
    return ("full house")
  }
  else if (is_flush(five_cards)) {
    return ("flush")
  }
  else if (is_straight(five_cards)) {
    return ("straight")
  }
  else if (is_three_of_a_kind(five_cards)) {
    return ("three of a kind")
  }
  else if (is_two_pair(five_cards)) {
    return ("two pair")
  }
  else if (is_pair(five_cards)) {
    return ("pair")
  }
  else {
    return ("nothing")
  }
}

before_draw()
#table(sapply(1:10000, FUN=before_draw()))
table(replicate(10000, before_draw()))
```

Calculated probabilities (number divided by 10,000): flush - 0.0016, four of a kind - 0.0001, full house - 0.0019, nothing - 0.5038, pair = 0.4188, straight - 0.0112, three of a kind - 0.0179, two pair - 0.0447.

Here is a break down in percentages: four of a kind - 0.01%, full house - 0.19%, flush - 0.16%, straight - 1.12%, three of a kind 0 1.79%, two pair - 4.47%, one pair - 41.88% and no pair - 50.38%.  These values are relatively close to the Wikipedia probabilities.  
## Test statements

```{r}
scoreActivity::score253(9)
```