# In-Class Programming Activity, Day 10

### *Hannah Sonsalla*

```{r}
library(ISLR)
data(Default, package = "ISLR")
```

## Task 1
Logistic Function
```{r}
logistic = function(x) {
  return(exp(x)/ (1+exp(x)))
}
```


## Task 2
Linear Combinations
```{r}
linear_combine = function(data, coefs){
  result <- 0
  for (nm in names(coefs)) {
    if (nm == "intercept") {
      result <- result + coefs[[nm]]
    } 
    else if (!nm %in% names(data)) {
      stop("Variable name in coef vector doesn't exist.")
    }
    else {
      result <- result + coefs[[nm]] * data[[nm]]
    }
  }  
  return (result)
}
```


##Task 3
Probabilities
```{r}
LL_logistic = function(data, coefs, outcome){
  ans1 = linear_combine(data, coefs)
  probs = logistic(ans1)
  likes = ifelse(outcome==1, probs, 1-probs)
  sum(log(likes))
}

#Test - should equal -2425.733
LL_logistic(data=Default, 
            coefs = c(intercept = 1, income = -0.0001), 
            outcome = Default$default == "Yes")
```

##Task 4
Optimize
```{r}

best_coefs <- optim(c(intercept = 1, income = 0.0001), LL_logistic, 
                    control=list(fnscale = -1), 
                    data = Default, 
                    outcome = Default$default == "Yes")
best_coefs
glm(default == "Yes" ~ income, data=Default, family="binomial")
```

Using an intercept of 1 and an income of 0.0001 results for coefs in optim() results in an intercept of -3.094342e+00 and income of -8.322106e-06.  When using the glm function, I obtained an intercept of -3.094e+00 and an income of -8.353e-06.  The intercepts of both of these are very close.  The incomes differ by about 0.031e-06.  
## Test statements

```{r}
scoreActivity::score253(10)
```