# In-Class Programming Activity, Day 15

### Hannah Sonsalla
Download & load Mona
```{r}
#download.file("http://tiny.cc/dcf/mona.rda", destfile = "mona.rda")
load("mona.rda")
```

## Task 1 - Overview
```{r}
X <- t(mona) - mean(mona[])
dim(X)
X_rand = matrix(rnorm(nrow(mona)*ncol(mona), mean = 0, sd = 1), ncol(mona), nrow(mona))
dim(X_rand)
X_corr = X_rand %*% chol(var(X))
dim(X_corr)
```

## Task 2 - Sparse beta
```{r}
beta = rep(0, 191)
sample_beta = sample(1:191, 16, replace = FALSE)

for (i in 1:16) {
  rand_num = sample(c(2, 5, -3, -4), 1)
  beta[sample_beta[i]] = rand_num
}
beta
```

## Task 3 - Output
```{r}
Y_pure = X %*% beta
var_Y_pure = var(Y_pure)
Y_real = X %*% beta + rnorm(nrow(X), mean = 0, sd = sqrt(0.1 * var_Y_pure))
```

## Task 4 - Least Squares
```{r}
beta_hat_pure = coef(lm(Y_pure ~ X-1))
plot(beta_hat_pure ~ beta)
```

```{r}
beta_hat_real = coef(lm(Y_real ~ X-1))
plot(beta_hat_real, beta)
```

Yes, we would be able to detect that beta is sparse because most of the data appears in a line.

## Task 4 - Lasso Estimator
```{r}
library(glmnet)
lasso_mod <- cv.glmnet(X, Y_real, alpha=1)
beta_lasso <- predict(lasso_mod, type = "coefficients", s = lasso_mod$lambda.min)
```

## Task 5 - Principal Components
```{r}
sing_vals_X = svd(X)$d
r_squared_X = (cumsum(sing_vals_X^2)) / (sum(sing_vals_X^2)) 
sing_vals_X_rand = svd(X_rand)$d
r_squared_X_rand = (cumsum(sing_vals_X_rand^2)) / (sum(sing_vals_X_rand^2)) 
sing_vals_X_corr = svd(X_corr)$d
r_squared_X_corr = (cumsum(sing_vals_X_corr^2)) / (sum(sing_vals_X_corr^2)) 

# Counts how many principal components necessary to reconstruct matrix with R^2 of 0.99%.
n99 = which(r_squared_X > 0.99)[[1]]
n99_rand = which(r_squared_X_rand > 0.99)[[1]]
n99_corr = which(r_squared_X_corr > 0.99)[[1]]

```


## Task 6 - Cross Validated R^2 function
```{r}
library(pls)
pcr.fit <- pcr(Y_real ~ X, scale = TRUE, validation="CV") 
r_squared_pcr.fit = R2(pcr.fit)
plot(r_squared_pcr.fit)
```

It appears that there are around 16 - 25 components necessary to get an r-squared of 0.85.


## Test statements

```{r}
scoreActivity::score253(15)
```