# In-Class Programming Activity, Day 16

### Hannah Sonsalla

## Task 1 - The Knots
```{r}
my_knots = function(x,k){
  qs = (1:k)/(1+k)
  #qs = seq(1/(k+1), 0.9999, by = 1/(k+1))
  quantile(x, probs=qs)
}

```

## Task 2 - The Basis Set
```{r}
spline_model_matrix = function(x, knot_locations){
  MM <- cbind(1, x, x^2, x^3)
  for (knot in knot_locations){
    new_vector = (x-knot)^3
    new_vector = ifelse(new_vector > 0, new_vector, 0)
    MM <- cbind(MM, new_vector)
  }
  return (MM)
}
```

## Task 3: Best linear combination
```{r}

fit_spline <- function(formula, k=2, data=parent.frame()) {
  y <- eval(formula[[2]], envir=data)
  x <- eval(formula[[3]], envir=data)
  knot_locations <- my_knots(x, k)
  MM <- spline_model_matrix(x, knot_locations)
  mod <- lm(y ~ MM - 1)
  # Modified to include RSE
  residual_degree_freedom = nrow(data) - ncol(MM)
  residuals = mod$residuals
  rse = sum(residuals^2) / residual_degree_freedom
  res <- list(coef = coef(mod), knots = knot_locations, cov = vcov(mod), rse = rse)
  class(res) <- "my_spline"
  
  return(res)
}
```


## Task 4: Predict Function
```{r}
predict.my_spline <- 
  function(mod, newx, level = 0.95,
           intervals=c("none", "confidence", "prediction")) {
  intervals <- match.arg(intervals)
  MM <- spline_model_matrix(newx, mod$knots)
  vals <- MM %*% mod$coef
  se <- sqrt(rowSums(MM %*% mod$cov * MM))
  rse = mod$rse
    if (intervals == "confidence") return(data.frame(upper = vals + 2* se, vals = vals, lower=vals - 2 * se))
  else if (intervals == "prediction") return(data.frame(upper = vals + 2 * sqrt(se^2 + rse^2), vals = vals, lower = vals - 2 * sqrt(se^2 + rse^2)))
  else return(vals) # for future use
}
```

Try function on Wage data from the ISLR package, finding a smooth model of wage versus age.  Tried with 4 knots.  
```{r}
library(ISLR)
data(Wage)
my_spline = fit_spline(wage~age, k=4, data = Wage)
num_seq = seq(20,80, length = 200)
y = predict(my_spline, newx = num_seq)
plot(num_seq,y, type = "l")
```

```{r}
#Confindence interval
confin_interval = predict(my_spline, newx = num_seq, intervals = "confidence")
plot(num_seq, confin_interval$vals, col = "black", type = "l", ylim = c(60, 125))
lines(num_seq, confin_interval$upper, col = "blue", lty = "dashed")
lines(num_seq, confin_interval$lower, col = "red", lty = "dashed")
```

```{r}
# Prediciton interval - need help
predict_interval = predict(my_spline, newx = num_seq, intervals = "predict")
```

Examine number of knots.
```{r}
num_seq = seq(20,80, length = 200)

my_spline1 = fit_spline(wage~age, k=2, data = Wage)
y1 = predict(my_spline1, newx = num_seq)

my_spline2 = fit_spline(wage~age, k=5, data = Wage)
y2 = predict(my_spline2, newx = num_seq)

my_spline3 = fit_spline(wage~age, k=8, data = Wage)
y3 = predict(my_spline3, newx = num_seq)

my_spline4 = fit_spline(wage~age, k=20, data = Wage)
y4 = predict(my_spline4, newx = num_seq)

par(mfrow=c(2,2))
plot(num_seq,y1, type = "l")
plot(num_seq,y2, type = "l")
plot(num_seq,y3, type = "l")
plot(num_seq,y4, type = "l")
```

Divide Wage into training and testing sets. Find the number of knots that minimizes the mean square error on the testing set.
```{r}
set.seed(1)
train = sample(1:dim(Wage)[1], dim(Wage)[1] / 2)
test = (-train)
Wage.train = Wage[train, ]
Wage.test = Wage[test, ]

# Using 3 knots
my_spline1 = fit_spline(wage~age, k=3, data = Wage.train)
pred = predict(my_spline1, newx = Wage.test$age)
mean((pred - Wage.test$age)^2)

# Using 10 knots
my_spline2 = fit_spline(wage~age, k=10, data= Wage.train)
pred2 = predict(my_spline2, newx = Wage.test$age)
mean((pred2 - Wage.test$age)^2)

# Using 25 knots
my_spline3 = fit_spline(wage~age, k=25, data = Wage.train)
pred3 = predict(my_spline3, newx = Wage.test$age)
mean((pred3 - Wage.test$age)^2)
```

To test this, I used 3 knots, 10 knots and 25 knots.  I found that 3 knots performed the best with the lowest MSE of 4863.558.

## Test statements
Note: These are the wrong tests for this activity.
```{r}
scoreActivity::score253(16)
```